---
title: "Data Analysis"
author: "Jieqi Tu"
date: "12/10/2020"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Import data
```{r import data}
HPV.data = readxl::read_excel("TCGA_HPV_mutation_Summary_042720 copy.xlsx")
```
First, we want extract genes: CD3e, CD3d, CD4, CD8a, CD8b, Foxp3, IFNg, IL2, PD1, PDL1, TP53, MYC, EGFR, ERBB2, ERBB3, EGF, NRG1 (Heregulin1/Neuregulin1), IL-6, TGFb1, FGF1, FGF2.


## Extract genes
```{r gene extraction}
HPV.data1 = 
  HPV.data %>% 
  filter(str_detect(name, c("CD3E")))
HPV.data2 = 
  HPV.data %>% 
  filter(str_detect(name, c("CD3D")))
HPV.data3 = 
  HPV.data %>% 
  filter(str_detect(name, c("CD4")))
HPV.data4 = 
  HPV.data %>% 
  filter(str_detect(name, c("CD8A")))
HPV.data5 = 
  HPV.data %>% 
  filter(str_detect(name, c("CD8B")))
HPV.data6 = 
  HPV.data %>% 
  filter(str_detect(name, c("FOXP3")))
HPV.data7 = 
  HPV.data %>% 
  filter(str_detect(name, c("IFNG")))
HPV.data8 = 
  HPV.data %>% 
  filter(str_detect(gene_name, c("COL8431")))
HPV.data9 = 
  HPV.data %>% 
  filter(str_detect(name, c("PD1"))) # not sure about the gene "PD1"
HPV.data10 = 
  HPV.data %>% 
  filter(str_detect(name, c("PDL"))) # not sure about the gene "PDL1"
HPV.data11 = 
  HPV.data %>% 
  filter(str_detect(name, c("TP53")))
HPV.data12 = 
  HPV.data %>% 
  filter(str_detect(name, c("MYC")))
HPV.data13 = 
  HPV.data %>% 
  filter(str_detect(name, c("EGFR")))
HPV.data14 = 
  HPV.data %>% 
  filter(str_detect(name, c("ERBB2")))
HPV.data15 = 
  HPV.data %>% 
  filter(str_detect(name, c("ERBB3")))
HPV.data16 = 
  HPV.data %>% 
  filter(str_detect(gene_name, c("COL5428")))
HPV.data17 = 
  HPV.data %>% 
  filter(str_detect(name, c("NRG1")))
HPV.data18 = 
  HPV.data %>% 
  filter(str_detect(name, c("IL6")))
HPV.data19 = 
  HPV.data %>% 
  filter(str_detect(name, c("TGFB1")))
HPV.data20 = 
  HPV.data %>% 
  filter(str_detect(gene_name, c("COL6387")))
HPV.data21 = 
  HPV.data %>% 
  filter(str_detect(gene_name, c("COL6392")))

HPV.data.new = rbind(HPV.data1, HPV.data2, HPV.data3, HPV.data4, HPV.data5, HPV.data6, HPV.data7,
                     HPV.data8, HPV.data9, HPV.data10, HPV.data11, HPV.data12, HPV.data13, HPV.data14,
                     HPV.data15, HPV.data16, HPV.data17, HPV.data18, HPV.data19, HPV.data20, HPV.data21)

# Write the results as an excel file
library(openxlsx)
write.xlsx(list(HPV.data1, HPV.data2, HPV.data3, HPV.data4, HPV.data5, HPV.data6, HPV.data7,
                HPV.data8, HPV.data9, HPV.data10, HPV.data11, HPV.data12, HPV.data13, HPV.data14,
                HPV.data15, HPV.data16, HPV.data17, HPV.data18, HPV.data19, HPV.data20, HPV.data21), "ExtracredGene.xlsx", sheetName = c("CD3e", "CD3d", "CD4", "CD8a", "CD8b", "Foxp3", "IFNg", "IL2", "PD1", "PDL1", "TP53", "MYC", "EGFR", "ERBB2", "ERBB3", "EGF", "NRG1",  "IL-6", "TGFb1", "FGF1", "FGF2"))

write.xlsx(HPV.data.new, "HPVdatanew.xlsx")
```

Genes are categorized in different sheets. 
Since we are not sure which genes are the targeted ones, we want reviewers to decide/filter what the correct genes should be.

Then we want to check whether there are other genes with p-values less than 0.05.

## Check genes with p-value < 0.05
```{r check genes}
# Select genes with p-values < 0.05
HPV.data.pval = 
  HPV.data %>% 
  filter(p_value <= 0.05)

nrow(HPV.data.pval)

# Write the results as an excel file
write.xlsx(HPV.data.pval, "Genes_smallP.xlsx")
```

Since there are 3609 genes having p-values less than 0.05. We cannot make a scatter point plot with all the gene names added on. 

```{r bar plot}
# Create bar plot with standard deviation
FGF1 = readxl::read_excel(path = "HPVdatanew.xlsx", sheet = "FGF1")
FGF1.plot = 
  FGF1 %>% 
  ggplot(aes(x = altered, y = mean)) + geom_bar(stat = "identity") + geom_errorbar(aes(x = altered, ymin = mean - StdDev, ymax = mean + StdDev)) +
  theme_bw() + labs(
    x = "FGF1",
    y = "Mean"
  )

TGFB1 = readxl::read_excel(path = "HPVdatanew.xlsx", sheet = "TGFB1")
TGFB1.plot = 
  TGFB1 %>% 
  ggplot(aes(x = altered, y = mean)) + geom_bar(stat = "identity") + geom_errorbar(aes(x = altered, ymin = mean - StdDev, ymax = mean + StdDev)) +
  theme_bw() + labs(
    x = "TGFB1",
    y = "Mean"
  )

FGF2 = readxl::read_excel(path = "HPVdatanew.xlsx", sheet = "FGF2")
FGF2.plot = 
  FGF2 %>% 
  ggplot(aes(x = altered, y = mean)) + geom_bar(stat = "identity") + geom_errorbar(aes(x = altered, ymin = mean - StdDev, ymax = mean + StdDev)) +
  theme_bw() + labs(
    x = "FGF2",
    y = "Mean"
  )

IL2 = readxl::read_excel(path = "HPVdatanew.xlsx", sheet = "IL2")
IL2.plot = 
  IL2 %>% 
  ggplot(aes(x = altered, y = mean)) + geom_bar(stat = "identity") + geom_errorbar(aes(x = altered, ymin = mean - StdDev, ymax = mean + StdDev)) +
  theme_bw() + labs(
    x = "IL2",
    y = "Mean"
  )

NRG1 = readxl::read_excel(path = "HPVdatanew.xlsx", sheet = "NRG1")
NRG1.plot = 
  NRG1 %>% 
  ggplot(aes(x = altered, y = mean)) + geom_bar(stat = "identity") + geom_errorbar(aes(x = altered, ymin = mean - StdDev, ymax = mean + StdDev)) +
  theme_bw() + labs(
    x = "NRG1",
    y = "Mean"
  )
library(ggpubr)
ggarrange(FGF1.plot, FGF2.plot, IL2.plot,
          NRG1.plot, TGFB1.plot,
          ncol = 2, nrow = 3,
          labels = c("1", "2", "3",
                     "4", "5"))
```

